---
title: "STATS 506 Final Project"
author: "Lingzhi Hao"
format:
  html: 
    code-fold: true
    embed-resources: true
  pdf: default
---

## **GitHub Repository:** <https://github.com/Lingzhi-Hao/STATS-506-Final-Project>

```{r}
set.seed(42)

library(dplyr)
library(e1071)
library(nortest)
library(ggplot2)
library(patchwork)
```

Set number of repetitions for Monte Carlo simulations, general sample size grid, and threshold of convergence.

```{r}
reps <- 5000
n_values <- c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 60, 70, 80, 100)

p_threshold <- 0.05
skew_threshold <- 0.1
kurt_threshold <- 0.2

```

1.  **Uniform distribution**

    Conduct Monte Carlo simulations to generate Z, and calculate the A2, p, Abs_Skew, and Abs_Kurtosis.

    ```{r}
    a_uni <- 0
    b_uni <- 1

    mu_uni <- (a_uni + b_uni) / 2
    sigma_uni <- sqrt((b_uni - a_uni)^2 / 12)

    results_uniform <- data.frame(
      n = integer(),
      A2 = numeric(),        # Anderson-Darling statistic
      p = numeric(),
      Abs_Skew = numeric(),  # Absolute Skewness
      Abs_Kurtosis = numeric() # Absolute Excess Kurtosis
    )

    for (n in n_values) {
      sample_means <- numeric(reps)

      for (i in 1:reps) {
        samples <- runif(n, min = a_uni, max = b_uni) 
        sample_means[i] <- mean(samples)
      }
      
      # Calculate the standardized Z-statistic
      Z <- (sample_means - mu_uni) / (sigma_uni / sqrt(n))
      
      
      # Anderson-Darling Statistic
      Z_sub <- sample(Z, 500)
      ad <- ad.test(Z_sub)
      ad_stat <- ad$statistic
      p_value <- ad$p.value
      
      # Absolute Skewness 
      skewness_val <- abs(e1071::skewness(Z))
      
      # Absolute Excess Kurtosis 
      kurtosis_val <- abs(e1071::kurtosis(Z)) 
      
      # Store results for the current n
      results_uniform <- rbind(results_uniform, data.frame(
        n = n,
        A2 = ad_stat, 
        p = p_value,
        Abs_Skew = skewness_val, 
        Abs_Kurtosis = kurtosis_val
      ))
    }

    # Print the resulting table
    print(results_uniform)
    ```

    Find the smallest sample size n\*.

    ```{r}
    n_star_uniform <- results_uniform %>%
      filter(
        p >= 0.05,
        Abs_Skew <= skew_threshold,
        Abs_Kurtosis <= kurt_threshold
      ) %>%
      arrange(n) %>%
      slice(1) %>%
      pull(n)

    n_star_uniform
    ```

    Regenerate Z using sample size n\*.

    ```{r}
    set.seed(42)

    sample_means_star <- numeric(reps)

    for (i in 1:reps) {
      samples <- runif(n_star_uniform, min = a_uni, max = b_uni)
      sample_means_star[i] <- mean(samples)
    }

    Z_star_uni <- (sample_means_star - mu_uni) / (sigma_uni / sqrt(n_star_uniform))

    ```

    Define a function to plot the histogram and Q-Q plot of Z.

    ```{r}
    plot_hist_qq <- function(Z_star, dist_name = "", n_star = NULL, bins = 40) {
      df_plot <- data.frame(Z = Z_star)

      title_suffix <- ""
      if (dist_name != "") title_suffix <- dist_name
      if (!is.null(n_star)) {
        title_suffix <- paste0(title_suffix, ifelse(title_suffix == "", "", ", "),
                               "n* = ", n_star)
      }
      if (title_suffix != "") title_suffix <- paste0(" (", title_suffix, ")")

      p_hist <- ggplot(df_plot, aes(x = Z)) +
        geom_histogram(aes(y = after_stat(density)),
                       bins = bins, fill = "skyblue", color = "white") +
        stat_function(fun = dnorm, color = "red", linewidth = 1) +
        labs(title = paste0("Histogram of Z", title_suffix),
             x = "Z", y = "Density") +
        theme_minimal()

      p_qq <- ggplot(df_plot, aes(sample = Z)) +
        stat_qq() +
        stat_qq_line(color = "red", linewidth = 1) +
        labs(title = paste0("Q–Q Plot of Z", title_suffix),
             x = "Theoretical Quantiles", y = "Sample Quantiles") +
        theme_minimal()

      p_hist + p_qq
    }

    ```

    Plot the histogram and Q-Q plot to see its closeness to normality.

    ```{r}
    plot_hist_qq(Z_star_uni, dist_name = "Uniform", n_star = n_star_uniform)

    ```

    Since the n\* calculated above is very small, refine the sample size grid to find a more precise n\*.

    ```{r}
    n_values_uni <- c(5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15)

    results_uniform_refine <- data.frame(
      n = integer(),
      A2 = numeric(),
      p = numeric(),
      Abs_Skew = numeric(),
      Abs_Kurtosis = numeric()
    )

    for (n in n_values_uni) {
      sample_means <- numeric(reps)

      for (i in 1:reps) {
        samples <- runif(n, min = a_uni, max = b_uni)
        sample_means[i] <- mean(samples)
      }

      Z <- (sample_means - mu_uni) / (sigma_uni / sqrt(n))

      Z_sub <- sample(Z, 500)
      ad <- ad.test(Z_sub)

      results_uniform_refine <- rbind(results_uniform_refine,
        data.frame(
          n = n,
          A2 = ad$statistic,
          p = ad$p.value,
          Abs_Skew = abs(skewness(Z)),
          Abs_Kurtosis = abs(kurtosis(Z))
      ))
    }

    print(results_uniform_refine)
    ```

    ```{r}
    n_star_uniform_refined <- results_uniform_refine %>%
      filter(
        p >= 0.05,
        Abs_Skew <= 0.1,
        Abs_Kurtosis <= 0.2
      ) %>%
      arrange(n) %>%
      slice(1) %>%
      pull(n)

    n_star_uniform_refined

    ```

    ```{r}
    set.seed(42)

    sample_means_star <- numeric(reps)

    for (i in 1:reps) {
      samples <- runif(n_star_uniform_refined, min = a_uni, max = b_uni)
      sample_means_star[i] <- mean(samples)
    }

    Z_star_uni_refined <- (sample_means_star - mu_uni) / (sigma_uni / sqrt(n_star_uniform_refined))

    plot_hist_qq(Z_star_uni_refined, dist_name = "Uniform", n_star = n_star_uniform_refined)
    ```

2.  **Chi-square distribution**

    ```{r}
    set.seed(42)
    df_chi <- 50 

    mu_chi <- df_chi          
    sigma_chi <- sqrt(df_chi * 2) 

    # Data frame to store results
    results_chisq <- data.frame(
      n = integer(),
      A2 = numeric(),        # Anderson-Darling statistic
      p = numeric(),
      Abs_Skew = numeric(),  # Absolute Skewness
      Abs_Kurtosis = numeric() # Absolute Excess Kurtosis
    )

    for (n in n_values) {
      sample_means <- numeric(reps)
      
      for (i in 1:reps) {
        samples <- rchisq(n, df = df_chi) 
        sample_means[i] <- mean(samples)
      }
      
      # Calculate the standardized Z-statistic)
      Z <- (sample_means - mu_chi) / (sigma_chi / sqrt(n))

      
      # Anderson-Darling Statistic
      Z_sub <- sample(Z, 500)
      ad <- ad.test(Z_sub)
      ad_stat <- ad$statistic
      p_value <- ad$p.value
      
      # Absolute Skewness
      skewness_val <- abs(e1071::skewness(Z))
      
      # Absolute Excess Kurtosis
      kurtosis_val <- abs(e1071::kurtosis(Z)) 
      
      # Store results for the current n
      results_chisq <- rbind(results_chisq, data.frame(
        n = n,
        A2 = ad_stat, 
        p = p_value,
        Abs_Skew = skewness_val, 
        Abs_Kurtosis = kurtosis_val
      ))
    }

    print(results_chisq)
    ```

    ```{r}
    n_star_chi <- results_chisq %>%
      filter(
        p >= 0.05,
        Abs_Skew <= skew_threshold,
        Abs_Kurtosis <= kurt_threshold
      ) %>%
      arrange(n) %>%
      slice(1) %>%
      pull(n)

    n_star_chi
    ```

    ```{r}
    set.seed(42)

    sample_means_star <- numeric(reps)

    for (i in 1:reps) {
      samples <- rchisq(n_star_chi, df = df_chi)
      sample_means_star[i] <- mean(samples)
    }

    Z_star_chi <- (sample_means_star - mu_chi) / (sigma_chi / sqrt(n_star_chi))

    plot_hist_qq(Z_star_chi, dist_name = "Chi-square", n_star = n_star_chi)
    ```

3.  **Log-normal distribution**

    ```{r}
    set.seed(42)

    mu_l <- 0
    sigma_l <- 0.3

    mu_log <- exp(mu_l + sigma_l^2 / 2)           
    sigma_log <- sqrt((exp(sigma_l^2) - 1) * exp(2 * mu_l + sigma_l^2))

    # Data frame to store results
    results_lognormal <- data.frame(
      n = integer(),
      A2 = numeric(),        # Anderson-Darling statistic
      p = numeric(),
      Abs_Skew = numeric(),  # Absolute Skewness
      Abs_Kurtosis = numeric() # Absolute Excess Kurtosis
    )

    for (n in n_values) {
      sample_means <- numeric(reps)
      
      for (i in 1:reps) {
        samples <- rlnorm(n, meanlog = mu_l, sdlog = sigma_l)
        sample_means[i] <- mean(samples)
      }
      
      # Calculate the standardized Z-statistic)
      Z <- (sample_means - mu_log) / (sigma_log / sqrt(n))

      Z_sub <- sample(Z, 500)
      ad <- ad.test(Z_sub)
      ad_stat <- ad$statistic
      p_value <- ad$p.value
      
      # Absolute Skewness
      skewness_val <- abs(e1071::skewness(Z))
      
      # Absolute Excess Kurtosis
      kurtosis_val <- abs(e1071::kurtosis(Z)) 
      
      # Store results for the current n
      results_lognormal <- rbind(results_lognormal, data.frame(
        n = n,
        A2 = ad_stat, 
        p = p_value,
        Abs_Skew = skewness_val, 
        Abs_Kurtosis = kurtosis_val
      ))
    }

    print(results_lognormal)
    ```

    ```{r}
    n_star_log <- results_lognormal %>%
      filter(
        p >= 0.05,
        Abs_Skew <= skew_threshold,
        Abs_Kurtosis <= kurt_threshold
      ) %>%
      arrange(n) %>%
      slice(1) %>%
      pull(n)

    n_star_log
    ```

    ```{r}
    set.seed(42)

    sample_means_star <- numeric(reps)

    for (i in 1:reps) {
      samples <- rlnorm(n_star_log, meanlog = mu_l, sdlog = sigma_l)
      sample_means_star[i] <- mean(samples)
    }

    Z_star_log <- (sample_means_star - mu_log) / (sigma_log / sqrt(n_star_log))

    plot_hist_qq(Z_star_chi, dist_name = "Log-normal", n_star = n_star_log)

    ```

4.  **Poisson distribution**

    ```{r}
    set.seed(42)
    lambda_p <- 15 

    mu_p <- lambda_p           
    sigma_p <- sqrt(lambda_p) 

    # Data frame to store results
    results_poisson <- data.frame(
      n = integer(),
      A2 = numeric(),        # Anderson-Darling statistic
      Abs_Skew = numeric(),  # Absolute Skewness
      p = numeric(),
      Abs_Kurtosis = numeric() # Absolute Excess Kurtosis
    )

    for (n in n_values) {
      sample_means <- numeric(reps)
      
      for (i in 1:reps) {
        samples <- rpois(n, lambda = lambda_p) 
        sample_means[i] <- mean(samples)
      }
      
      # Calculate the standardized Z-statistic)
      Z <- (sample_means - mu_p) / (sigma_p / sqrt(n))

      
      # Anderson-Darling Statistic
      Z_sub <- sample(Z, 500)
      ad <- ad.test(Z_sub)
      ad_stat <- ad$statistic
      p_value <- ad$p.value
      
      # Absolute Skewness
      skewness_val <- abs(e1071::skewness(Z))
      
      # Absolute Excess Kurtosis
      kurtosis_val <- abs(e1071::kurtosis(Z)) 
      
      # Store results for the current n
      results_poisson <- rbind(results_poisson, data.frame(
        n = n,
        A2 = ad_stat, 
        p = p_value,
        Abs_Skew = skewness_val, 
        Abs_Kurtosis = kurtosis_val
      ))
    }

    print(results_poisson)
    ```

    ```{r}
    n_star_p <- results_poisson %>%
      filter(
        p >= 0.05,
        Abs_Skew <= skew_threshold,
        Abs_Kurtosis <= kurt_threshold
      ) %>%
      arrange(n) %>%
      slice(1) %>%
      pull(n)

    n_star_p
    ```

    ```{r}
    set.seed(42)

    sample_means_star <- numeric(reps)

    for (i in 1:reps) {
      samples <- rpois(n_star_p, lambda = lambda_p)
      sample_means_star[i] <- mean(samples)
    }

    Z_star_p <- (sample_means_star - mu_p) / (sigma_p / sqrt(n_star_p))

    plot_hist_qq(Z_star_p, dist_name = "Poisson", n_star = n_star_p)
    ```

5.  **t-distribution**

    ```{r}
    set.seed(42)
    df_t <- 10 

    mu_t <- 0           
    sigma_t <- sqrt(df_t / (df_t - 2)) 

    # Data frame to store results
    results_t <- data.frame(
      n = integer(),
      A2 = numeric(),        # Anderson-Darling statistic
      p = numeric(),
      Abs_Skew = numeric(),  # Absolute Skewness
      Abs_Kurtosis = numeric() # Absolute Excess Kurtosis
    )

    for (n in n_values) {
      sample_means <- numeric(reps)
      
      for (i in 1:reps) {
        samples <- rt(n, df = df_t)
        sample_means[i] <- mean(samples)
      }
      
      # Calculate the standardized Z-statistic)
      Z <- (sample_means - mu_t) / (sigma_t / sqrt(n))

      
      # Anderson-Darling Statistic
      Z_sub <- sample(Z, 500)
      ad <- ad.test(Z_sub)
      ad_stat <- ad$statistic
      p_value <- ad$p.value
      
      # Absolute Skewness
      skewness_val <- abs(e1071::skewness(Z))
      
      # Absolute Excess Kurtosis
      kurtosis_val <- abs(e1071::kurtosis(Z)) 
      
      # Store results for the current n
      results_t <- rbind(results_t, data.frame(
        n = n,
        A2 = ad_stat,     
        p = p_value,
        Abs_Skew = skewness_val, 
        Abs_Kurtosis = kurtosis_val
      ))
    }

    print(results_t)
    ```

    ```{r}
    n_star_t <- results_t %>%
      filter(
        p >= 0.05,
        Abs_Skew <= skew_threshold,
        Abs_Kurtosis <= kurt_threshold
      ) %>%
      arrange(n) %>%
      slice(1) %>%
      pull(n)

    n_star_t
    ```

    ```{r}
    set.seed(42)

    sample_means_star <- numeric(reps)

    for (i in 1:reps) {
      samples <- rt(n_star_t, df = df_t)
      sample_means_star[i] <- mean(samples)
    }

    Z_star_t <- (sample_means_star - mu_t) / (sigma_t / sqrt(n_star_t))

    plot_hist_qq(Z_star_t, dist_name = "Student's t", n_star = n_star_t)
    ```

6.  **Bernoulli distribution**

    ```{r}
    set.seed(42)
    n_values_b <- c(5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 60, 70, 80, 100, 120, 140, 160, 180, 200)
    p_b <- 0.5 

    mu_b <- p_b           
    sigma_b <- sqrt(p_b * (1 - p_b)) 

    # Data frame to store results
    results_bern <- data.frame(
      n = integer(),
      A2 = numeric(),        # Anderson-Darling statistic
      p = numeric(),
      Abs_Skew = numeric(),  # Absolute Skewness
      Abs_Kurtosis = numeric() # Absolute Excess Kurtosis
    )

    for (n in n_values_b) {
      sample_means <- numeric(reps)
      
      for (i in 1:reps) {
        samples <- rbinom(n, size = 1, prob = p_b) 
        sample_means[i] <- mean(samples)
      }
      
      # Calculate the standardized Z-statistic)
      Z <- (sample_means - mu_b) / (sigma_b / sqrt(n))

      
      # Anderson-Darling Statistic
      Z_sub <- sample(Z, 500)
      ad <- ad.test(Z_sub)
      ad_stat <- ad$statistic
      p_value <- ad$p.value
      
      # Absolute Skewness
      skewness_val <- abs(e1071::skewness(Z))
      
      # Absolute Excess Kurtosis
      kurtosis_val <- abs(e1071::kurtosis(Z)) 
      
      # Store results for the current n
      results_bern <- rbind(results_bern, data.frame(
        n = n,
        A2 = ad_stat,
        p = p_value,
        Abs_Skew = skewness_val, 
        Abs_Kurtosis = kurtosis_val
      ))
    }

    print(results_bern)
    ```

    ```{r}
    n_star_b <- results_bern %>%
      filter(
        p >= 0.05,
        Abs_Skew <= skew_threshold,
        Abs_Kurtosis <= kurt_threshold
      ) %>%
      arrange(n) %>%
      slice(1) %>%
      pull(n)

    n_star_b
    ```

    ```{r}
    set.seed(42)

    sample_means_star <- numeric(reps)

    for (i in 1:reps) {
      samples <- rbinom(n_star_b, size = 1, prob = p_b)
      sample_means_star[i] <- mean(samples)
    }

    Z_star_b <- (sample_means_star - mu_b) / (sigma_b / sqrt(n_star_b))

    plot_hist_qq(Z_star_b, dist_name = "Bernoulli", n_star = n_star_b)


    ```

7.  **Mixed-normal distribution**

    ```{r}
    set.seed(42)
    n_values_m = c(100, 120, 150, 180, 200, 250, 300)
    w1 <- 0.9
    w2 <- 1 - w1
    mu1 <- 2
    mu2 <- -2
    sd1 <- 1
    sd2 <- 1

    mu_m <- w1 * mu1 + w2 * mu2
    sigma_m <- sqrt(w1 * (mu1^2 + sd1^2) + w2 * (mu2^2 + sd2^2) - mu_m^2)

    rmixnorm <- function(n, w1, mu1, sd1, w2, mu2, sd2) {
      component <- sample(c(1, 2), size = n, replace = TRUE, prob = c(w1, w2))
      samples <- numeric(n)
      n1 <- sum(component == 1)
      samples[component == 1] <- rnorm(n1, mean = mu1, sd = sd1)
      n2 <- sum(component == 2)
      samples[component == 2] <- rnorm(n2, mean = mu2, sd = sd2)
      
      return(samples)
    }


    # Data frame to store results
    results_mixnorm <- data.frame(
      n = integer(),
      A2 = numeric(),        # Anderson-Darling statistic
      p = numeric(),
      Abs_Skew = numeric(),  # Absolute Skewness
      Abs_Kurtosis = numeric() # Absolute Excess Kurtosis
    )

    for (n in n_values_m) {
      sample_means <- numeric(reps)
      
      for (i in 1:reps) {
        samples <- rmixnorm(n, w1, mu1, sd1, w2, mu2, sd2) 
        sample_means[i] <- mean(samples)
      }
      
      # Calculate the standardized Z-statistic)
      Z <- (sample_means - mu_m) / (sigma_m / sqrt(n))

      
      # Anderson-Darling Statistic
      Z_sub <- sample(Z, 500)
      ad <- ad.test(Z_sub)
      ad_stat <- ad$statistic
      p_value <- ad$p.value
      
      # Absolute Skewness
      skewness_val <- abs(e1071::skewness(Z))
      
      # Absolute Excess Kurtosis
      kurtosis_val <- abs(e1071::kurtosis(Z)) 
      
      # Store results for the current n
      results_mixnorm <- rbind(results_mixnorm, data.frame(
        n = n,
        A2 = ad_stat, 
        p = p_value,
        Abs_Skew = skewness_val, 
        Abs_Kurtosis = kurtosis_val
      ))
    }

    print(results_mixnorm)
    ```

    ```{r}
    n_star_m <- results_mixnorm %>%
      filter(
        p >= 0.05,
        Abs_Skew <= skew_threshold,
        Abs_Kurtosis <= kurt_threshold
      ) %>%
      arrange(n) %>%
      slice(1) %>%
      pull(n)

    n_star_m
    ```

    ```{r}
    set.seed(42)

    sample_means_star <- numeric(reps)

    for (i in 1:reps) {
      samples <- rmixnorm(n_star_m, w1, mu1, sd1, w2, mu2, sd2)
      sample_means_star[i] <- mean(samples)
    }

    Z_star_m <- (sample_means_star - mu_m) / (sigma_m / sqrt(n_star_m))

    plot_hist_qq(Z_star_m, dist_name = "Mixed-normal", n_star = n_star_m)
    ```
